---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# kvkapiR

<!-- badges: start -->

`r lifecycle::badge('experimental')`
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of kvkapiR is to provides a convenient R programming language interface to the Dutch Chamber of Commerce (KvK) Search API. This package is built using the `httr2` package and follows best practices for wrapping APIs in R, as outlined in the httr2 documentation (<https://httr2.r-lib.org/articles/wrapping-apis.html>). It simplifies authentication, request handling, and response parsing when interacting with the KvK API.

Currently, the package only provides access to the [KvK search API](https://developers.kvk.nl/apis/zoeken). Additional APIs are available; more details can be found on the  [on the developers website of the KvK](https://developers.kvk.nl/apis).

## Installation

You can install the development version of `kvkapiR` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("coeneisma/kvkapiR")
```

To install the latest development version:

``` r
# install.packages("devtools")
devtools::install_github("coeneisma/kvkapiR", ref = "development")
```

## Example

Here are basic examples demonstrating how to use the `kvk_search()` function to retrieve KvK registrations:

```{r example}
library(kvkapiR)

koudum <- kvk_search(plaats = "Koudum")
koudum
```

You can also combine multiple search parameters:

```{r}
library(kvkapiR)

rotterdam <- kvk_search(naam = "huisartsen", plaats = "Rotterdam")
rotterdam
```

See `vignette("kvkapiR")` for more examples and information about the usage of this package.
