---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# kvkapiR

<!-- badges: start -->

`r lifecycle::badge('experimental')`
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of kvkapiR is to provide a convenient R programming language interface to the Dutch Chamber of Commerce (KvK) APIs. This package is built using the httr2 package and follows best practices for wrapping APIs in R, as outlined in the [httr2 documentation](https://httr2.r-lib.org/articles/wrapping-apis.html). It simplifies authentication, request handling, and response parsing when interacting with the KvK API.

Currently, the package provides access to the [following KvK APIs](https://developers.kvk.nl/documentation#available-apis):

* KvK Search API
* KvK Basisprofiel API
* KvK Vestigingsprofiel API
* KvK Naamgeving API

More details can be found [on the developers website of the KvK](https://developers.kvk.nl/apis).

## Installation

You can install the development version of `kvkapiR` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("coeneisma/kvkapiR")
```

To install the latest development version:

``` r
# install.packages("devtools")
devtools::install_github("coeneisma/kvkapiR", ref = "development")
```

## Example

Here are basic examples demonstrating how to use the `kvk_search()` function to retrieve KvK registrations:

```{r example}
library(kvkapiR)

koudum <- kvk_search(plaats = "Koudum")
koudum
```

You can also combine multiple search parameters:

```{r}
library(kvkapiR)

rotterdam <- kvk_search(naam = "huisartsen", plaats = "Rotterdam")
rotterdam
```

See `vignette("kvkapiR")` for more examples and information about the usage of this package.
