---
title: "kvkapiR"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

# Check if API key is available for conditional chunk execution
has_api_key <- nzchar(Sys.getenv("KVK_API_KEY"))
```

The goal of kvkapiR is to provide a convenient R programming language interface to the Dutch Chamber of Commerce (KvK) APIs. This package is built using the httr2 package and follows best practices for wrapping APIs in R, as outlined in the [httr2 documentation](https://httr2.r-lib.org/articles/wrapping-apis.html). It simplifies authentication, request handling, and response parsing when interacting with the KvK API.

The package provides access to [all the KvK APIs](https://developers.kvk.nl/documentation#available-apis):

* KvK Search API
* KvK Basisprofiel API
* KvK Vestigingsprofiel API
* KvK Naamgeving API

More details on the API can be found [on the developers website of the KvK](https://developers.kvk.nl/apis).

## Loading the package

```{r}
library(kvkapiR)
```

## API keys

```{r, echo = FALSE, results = 'asis'}
if (!has_api_key) {
  cat("**Note:** No API key was found. The production API examples below will not be executed. ")
  cat("To run all examples, please set your API key using `kvk_set_api_key()` or the test environment examples will still work.")
}
```

### Optaining an API key

To access the KvK API, you need an API key. You can apply for an API key and learn about different KvK APIs on the [KvK Developer Portal](https://developers.kvk.nl/apply-for-apis?step=api-overview).

### Temporary API Key (Current Session Only)

If you only need an API key for the current session, use the `kvk_set_api_key()` function. 

``` r
kvk_set_api_key(api_key = "your_api_key_here")
```

**âš ï¸ Note:** This only sets the key for the current R session. The key will be lost once you restart R.

### Persistent API Key (Stored in `.Renviron`)

If you want to store the API key permanently, use the `kvk_store_api_key()` function. This saves the key in the `.Renviron` file, making it persist across R sessions.

``` r
kvk_store_api_key(api_key = "your_api_key_here")
```
To update an existing API key, set the `overwrite` parameter to `TRUE`:

``` r 
kvk_store_api_key(api_key = "your_new_api_key", overwrite = TRUE)
```
**âš ï¸ Important:** You must restart R for the changes to take effect.

**âš ï¸ Warning:** Only use this method if you can safely store your API key on your system.

### Testing Environmet

KvK provides a test environment where you can experiment with the API before using it in production. More information is available on the [KvK documentation page](https://developers.kvk.nl/documentation/testing).

All functions in `kvkapiR` support testing mode via the `test_environment` argument, which is set to `FALSE` by default.

**ðŸ¦† Tip:** The test environment contains fictional data based on characters from the Dutch comic magazine "Donald Duck". Try searching for names like "Donald" and "Dagobert" to see how the API works with sample data.

## Search the KvK API

The `kvk_search()` function is wrapper for the KvK Search API. It automatically paginates results (up to 1,000 records).

Below are examples demonstrating how to perform searches and handle API responses.

### Example 1: single variable, less then 1.000 records

Lets search for KvK registrations in [the village of Koudum](https://nl.wikipedia.org/wiki/Koudum). Since this is a small village with fewer than *1.000* registered businesses (the API limit), all records are retrieved in one request.

```{r, eval = has_api_key}
koudum <- kvk_search(plaats = "Koudum")
koudum
```


### Example 2: single variable, more then 1.000 records

The [city of Rotterdam](https://nl.wikipedia.org/wiki/Rotterdam) is significantly larger than Koudum. Searching for all registrations in Rotterdam will return more than 1,000 records, exceeding the API limit. 
When this happens, a warning is displayed:

```{r, eval = has_api_key}
rotterdam <- kvk_search(plaats = "Rotterdam")
rotterdam
```

### Example 3: combining multiple search terms in variables

You can combine multiple search terms in any parameter to refine your search. The API treats multiple words within a parameter as an AND operation by default.

```{r, eval = has_api_key}
# Search for businesses with both "snackbar" and "hoekje" in their name
snackbar_hoekje <- kvk_search(naam = "snackbar hoekje")
snackbar_hoekje
```



### Example 4: Searching with Multiple Criteria

You can combine multiple search criteria in a query.

For example, let's search for businesses in Utrecht where the name contains both `snackbar` and `hoek`:

```{r, eval = has_api_key}
snackbar <- kvk_search(plaats = "Utrecht", naam = "snackbar")
snackbar
```

### Example 5: search on adress

It is possible to search for KvK registrations using a specific address. To do this, you need both the postal code and house number.
For example, are there any business registrered at [this adress](https://maps.app.goo.gl/KJhYVDH3fV8cxWJ2A)? 

```{r, eval = has_api_key}
secret_adress <- kvk_search(postcode = "2594BD", huisnummer = "10")
secret_adress
```

## Retrieving Profiel Data

The package has the abbility to retrieve three types of *profielen* from the API:

* basisprofiel
* vestigingsprofiel
* naamgeving

For these examples we will use the test invironment.

```{r}
donald <- kvk_search(naam = "Donald", test_environment = TRUE)
```


### Basisprofiel

```{r}
kvk_get_basisprofiel(kvkNummer = donald$kvkNummer[1], test_environment = TRUE)
```


### Vestigingsprofiel


```{r}
kvk_get_vestigingsprofiel(vestigingsnummer = donald$vestigingsnummer[1], test_environment = TRUE)
```

### Naamgeving

```{r}
kvk_get_naamgeving(kvkNummer = donald$kvkNummer[1], test_environment = TRUE)
```


## Conclusion

The `kvkapiR` package provides a simple yet powerful interface for searching the KvK API in R. With built-in authentication handling, test environment support, and flexible query options, it streamlines access to Dutch business registration data.

For more details, refer to the official [KvK API documentation](https://developers.kvk.nl/apis).
