---
title: "kvkapiR"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


The kvkapiR package provides a user-friendly interface to the Dutch Chamber of Commerce (KvK) Search API within the R programming environment. Built using the `httr2` package, it adheres to best practices for API wrappers in R, as outlined in the [httr2 documentation](https://httr2.r-lib.org/articles/wrapping-apis.html).

This package simplifies authentication, request handling, and response parsing, making it easier to interact with the KvK API.

Currently, kvkapiR only supports the [KvK search API](https://developers.kvk.nl/apis/zoeken). Additional KvK APIs are available, and more details can be found [on the developers website of the KvK](https://developers.kvk.nl/apis).


## Loading the package

```{r}
library(kvkapiR)
```

## Optaining an API key

To access the KvK API, you need an API key. You can apply for an API key and learn about different KvK APIs on the [KvK Developer Portal](https://developers.kvk.nl/apply-for-apis?step=api-overview).

### Testing Environmet

KvK provides a test environment where you can experiment with the API before using it in production. More information is available on the [KvK documentation page](https://developers.kvk.nl/apply-for-apis?step=api-overview).

All functions in `kvkapiR` support testing mode via the `test_environment` argument, which is set to `FALSE` by default.

## Setting Up the API Key in R

### Temporary API Key (Current Session Only)

If you only need the API key for the current session, use the `kvk_set_api_key()``` function.

``` r
kvk_set_api_key("put_your_api_key_here")
```
**⚠️ Note:** This only sets the key for the current R session. The key will be lost once you restart R.

### Persistent API Key (Stored in `.Renviron`)

If you want to store the API key permanently, use the `kvk_store_api_key()` function. This saves the key in the `.Renviron` file, making it persist across R sessions.

``` r
kvk_store_api_key("put_your_api_key_here")
```
**⚠️ Warning:** Only use this method if you can safely store your API key on your system.

## Using the KvK API with `kvkapiR`

The `kvkapiR` package acts as a simple wrapper around the KvK API. Below are examples demonstrating how to perform searches and handle API responses.

### Example 1: single variable, less then 1.000 records

Lets search for KvK registrations in [the village of Koudum](https://nl.wikipedia.org/wiki/Koudum). Since this is a small village with fewer than *1.000* registered businesses (the API limit), all records are retrieved in one request.

```{r}
koudum <- kvk_search(plaats = "Koudum")
koudum
```

### Example 2: single variable, more then 1.000 records

The [city of Rotterdam](https://nl.wikipedia.org/wiki/Rotterdam) is significantly larger than Koudum. earching for all registrations in Rotterdam will return more than 1,000 records, exceeding the API limit. 
When this happens, a warning is displayed:

```{r}
rotterdam <- kvk_search(plaats = "Rotterdam")
rotterdam
```

### Example 3: Searching with Multiple Criteria

You can combine multiple search criteria in a query.

For example, let's search for businesses in Utrecht where the name contains both `snackbar` and `hoek`:

```{r}
snackbar <- kvk_search(plaats = "Utrecht", naam = "snackbar")
snackbar
```

### Example 4: search on adress

It is possible to search for KvK registrations using a specific address. To do this, you need both the postal code and house number.
For example, are there any business registrered at [this adress](https://maps.app.goo.gl/KJhYVDH3fV8cxWJ2A)? 

```{r}
secret_adress <- kvk_search(postcode = "2594BD", huisnummer = "10")
secret_adress
```

## Conclusion

The `kvkapiR` package provides a simple yet powerful interface for searching the KvK API in R. With built-in authentication handling, test environment support, and flexible query options, it streamlines access to Dutch business registration data.

For more details, refer to the official [KvK API documentation](https://developers.kvk.nl/apis).
